---
title: <center> <h1> Homework 4 </h1> </center>
mainfont: Arial
author: "Boxiang Jia"
date: December 4, 2018
output: html_document
runtime: shiny 
fontsize: 12pt
geometry: margin=1in
self_contained: yes
---

## Exercise 5.8 Problem 2
we first create a function to see if a matix has its inverse matrix:
```{r}
f <- function(m) class(try(solve(m),silent=T))=="matrix"
```
now create the matrix X and probability p
Set
```{r}
X <- rbind(c(1, -1/4), c(-1/4, 1))  
p <- c(0.5, 0)
```
To get the linear hessian $XtX$ be well condition, 
```{r}
A = t(X)%*%X
f(A)
```
According to the book, the logistic variation of Hessian matrix is 
$$X'diag(p(1-p))X$$
```{r}
B <- t(X)%*%diag(p*(1-p))%*%X
f(B)
```
In this way，the linear Hessian (XtX) is well-conditioned, the function we created to check if a matrix has a inverse returns a "True", but the logistic variation is not, which does not have a inverse matrix.

##Exercise 5.8 Problem 4
Accordign to the book and reference to the help and code by Yingnan Lv, 
$$
l'(y) = l'(y) - \lambda||\beta||_2^2.
$$
We take the derivative on $\beta$ first, then we will ahve 
$$
\Delta_{\beta}l'(y) = X^t(y-Ey) - 2\lambda\beta.
$$
then we take the second derivative, and we will have 

$$
h(l') = \frac{\delta^2 l'(y)}{\delta \beta_k \delta \beta_j}
= - X_t \cdot \text{diag}(Var(y)) \cdot X - 2\lambda I_{p}
$$

then we import the derivative and hessian matrix to get the new beta and find out the relation between V and XtX

$$
\begin{aligned}
\beta^{(k+1)}&=\beta^{(k)}-H^{-1}(l)(\beta^{(k)})\triangle_{\beta}l(y)(\beta^{(k)})\\
&=\beta^{(k)}+[X^t\text{diag}(var(y^{(k)}))X+2\lambda I_{p}]^{-1}(X^t(y-{E}y^{(k)})-2\lambda\beta)\\
&=V^{-1}X^tW(X\beta^{k}+W^{-1}(y-{E}y^{k}))\\
&=\frac {X^tWz}{(X^tWX+2\lambda I_{p})}
\end{aligned}
$$
hence we have  $V = X^t W X + 2 \lambda I_{p}$
and then we change it in the casl_glm_irwls function as below.
```{r}
#modify the function 
casl_glm_irwls <- function(X, y, family, maxit=25, tol=1e-10,lamda){
  beta <- rep(0,ncol(X))
  for(j in seq_len(maxit)){
    b_old <- beta
    eta <- X %*% beta
  mu <- family$linkinv(eta)
  mu_p <- family$mu.eta(eta)
  z <- eta + (y - mu) / mu_p
  W <- as.numeric(mu_p^2 / family$variance(mu))
  XtX <- crossprod(X, diag(W) %*% X)
  V <-  XtX + 2*diag(rep(lambda, dim(XtX)[1]))
  Xtz <- crossprod(X, W * z)
  beta <- solve(XtX, Xtz)
    if(sqrt(crossprod(beta - b_old)) < tol) break
  }
beta 
}
#test the function
n <- 1000; p <- 3
beta <- c(0.2, 2, 1)
X <- cbind(1, matrix(rnorm(n * (p- 1)), ncol = p - 1))
mu <- 1 - pcauchy(X %*% beta)
y <- as.numeric(runif(n) > mu)
beta <- casl_glm_irwls_glm(X, y,family = binomial(link = "cauchit"),lamda = 0.5)
beta
```
